---
title: "Lab 1 - Data visualization"
author: "[Ann Chang]"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: inline
---

## Load Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(viridis)

```

## Exercise 1

```{r}
#| label: histogram of population density

ggplot(data = midwest, 
       aes(x = popdensity)) +
  geom_histogram(binwidth = 10000) +
  labs(
    x = "Population Density (person/unit area)",
    y = "Frequency",
    title = 
      "    Population Densities of Midwest Counties
               (2000 US Census)"
    )

```
- The distribution of the data seem to be right-skewed.
- There seem to be some outliers of relatively large population densities, such
as the incidences beyond 50000 people/unit area. 

## Exercise 2

(Type your answer to Exercise 2 here. Add code chunks as needed. Don't forget to label your code chunk. Do not use spaces in code chunk labels.)

## Exercise 3

(Type your answer to Exercise 3 here. Add code chunks as needed. Don't forget to label your code chunk. Do not use spaces in code chunk labels.)

## Exercise 4

## Exercise 5

## Exercise 6

## Exercise 7
