---
title: "Lab 1 - Data visualization"
author: "[Ann Chang]"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: inline
---

## Load Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(viridis)

```

## Exercise 1

```{r}
#| label: ex-1 histogram of population density
#| message: false

ggplot(data = midwest, 
       aes(x = popdensity)) +
  geom_histogram(binwidth = 10000) +
  labs(
    x = "Population Density (person/unit area)",
    y = "Frequency",
    title = 
      "    Population Densities of Midwest Counties
               (2000 US Census)"
    )

```
- The distribution of the data seem to be right-skewed.
- There seem to be some outliers of relatively large population densities, such
as the incidences beyond 50000 people/unit area. 

## Exercise 2


```{r}
#| label: ex-2 scatterplot of education vs poverty level
#| message: false

ggplot(data = midwest, 
       aes(x = percollege,
           y = percbelowpoverty,
           color = state)) +
  geom_point(alpha = 0.6) +
  labs(
    x = "Percentage of People with a College Degree",
    y = "Percentage Below Poverty",
    title = "Relationship Between Education Level and Poverty Level",
    color = "State"
  ) +
  scale_color_viridis_d()

```

## Exercise 3

There seems to be a negative correlation between the percentage of people below the poverty line and the percentage of people with a college degree. Illinois 
seems to have relatively more outlying counties where the percentage of below poverty individuals is larger than what would be expected from the percentage of people with a college degree. This is also seen in some counties in Ohio. The counties in Wisconsin are relatively clustered closely together, with no counties' percentage of individuals below poverty exceeding above 23%. 

## Exercise 4

```{r}
#| label: ex-4 scatterplots separated by state
#| message: false

state.labs <- c("Illinois", "Indiana", "Michigan", "Ohio", "Wisconsin")
names(state.labs) <- c("IL", "IN", "MI", "OH", "WI")

ggplot(data = midwest, 
       aes(x = percollege,
           y = percbelowpoverty,
           color = state)) +
  geom_point(alpha = 0.6) +
  facet_wrap(~state, 
             labeller = labeller(state = state.labs)
             ) +
  geom_smooth(se = FALSE) +
  scale_color_viridis_d() +
  labs(
    x = "Percentage of Individuals with a College Degree",
    y = "Percentage of Individuals Below Poverty ",
    title = "Relationship Between College Degree Aquisition and Poverty Level,
                            by State",
    color = "State"
    ) +
  scale_color_discrete(labels = state.labs)
```
The faceted scatterplot is much more preferable than the original scatterplot
from Exercise 2 because it is much easier to see the relationship between these
two percentages for each individual state. This relationship for each state was
much harder to see in the plot from Exercise 2 because the points were too crowded together to be able to separate the states, even with the differential coloring. 

## Exercise 5

## Exercise 6

## Exercise 7
